|

### 2020-05-25 paper/Manuscript.Rmd
Further improvements

abstract: |
  Technological advances enable researchers to more easily conduct reliable,
  cumulative, collaborative science. Capitalizing on these advances has the
  potential to accelerate scientific progress. However, adopting open science
  principles can be challenging and costly, because doing so requires
  substantial conceptual education and training in the use of new tools. The 
  goal of this paper is to remove these obstacles. We introduce the Workflow for
  Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
  procedure that researchers can follow to make an entire project open and
  reproducible. It is based on best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies. To further facilitate the transition, the
  WORCS principles are implemented in an R-package, which offers an RStudio 
  project template that is, effectively, a "one-click solution" for R users. 
  This paper introduces the workflow, discusses how it meets different 
  requirements of open science, and provides examples of the implementation of
  `worcs` in R.

@@

The (social) sciences are arguably past the tipping point of a paradigm shift
towards open science.
Several immediate causes rallied support for this transition, including
highly publicized cases of scientific fraud [as discussed by
@leveltFailingScienceFraudulent2012], increasing awareness of questionable
research practices and their consequences for the credibility of findings [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis
[@shroutPsychologyScienceKnowledge2018].
However, open science should not be seen as a cure (or punishment) for this
crisis. As the late dr. Jonathan Tennant put it: "Open science is just good science"
[@tennantOpenScienceJust2018].
Moreover, open science creates opportunities for researchers to more easily
conduct reliable, cumulative, and 
collaborative science [@nosekScientificUtopiaOpening2012;
@adolphOpenBehavioralScience2012]. Open science also
promotes inclusivity, because it removes barriers for participation.
Capitalizing on these advances has the potential to accellerate scientific
progress [see also @coyneReplicationInitiativesWill2016].

Many individual researchers are motivated to adopt best practices for open
science and enjoy these benefits. And yet, the question of "how" to 
adopt open science practices can be daunting. Making the transition
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by providing a simple workflow 
that meets most requirements open science, is based on best practices, and 
is supported by user-friendly software in R: The Worfkflow for Open Reproducible 
Code in Science (WORCS). WORCS is a lightweight approach to open science and
computational reproducibility. The workflow (Figure <!--\@ref(fig:workflow)-->) 
constitutes a step-by-step procedure that researchers can follow to make a 
research project open and reproducible. WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them. It can also be used in the absence of top-down support or guidelines
to produce work in accordance with best practices.
Although the principles underlying
the WORCS procedure are universal, they have been implemented for R users in the
R-package `worcs`, which offers an RStudio project template that is, effectively,
a "one click solution". This paper introduces the WORCS workflow, discusses how
it meets best practices for open science, and illustrates the use of the `worcs`
package.

<!-- AL: I second LB's comment somewhere below that the overall WORCS workflow should be introduced much earlier, maybe already here. A diagrammatic overview/flowchart of the principal steps would be great. It should catch the reader's eye and give an immediate idea of what WORCS is about. Now it is indeed quite buried in the details that follow. -->

<!--
thereby lowering the threshold for adoption.
, and still ensures reproducibility in most circumstances. 
 It is easy to implement, thereby lowering the threshold

* Grass-roots
* open science as opportunity
    - Easier work
    - Less mistakes
    - Better collaboration
    - Cumulative science
    - Error correction

Benefits to individual scientists: Automation (after a learning curve), thus more productive, avoid mistakes, join projects, clone projects and make them your own.
Collective benefits: Cumulative science, -->
<!-- AL: These "value propositions" would also be good to discuss already in the introduction. Directly after the introduction of the WORCS overview would make sense, I think. -->

<!-- AL: The remainder of the paper is structured as follows. ... Or some variation of this obligatory end-of-intro paragraph. As the paper touches on a lot of aspects, from principal to very practical, it would be very useful to provide some guidance here. 
CJ: I was always taught that you should not explain your structure in the paper - you should just structure the paper. If the structure is unclear, make it better. -->
<!--openness, transparency, and reproducible science -->

@@

Although Open Science is advocated by many, it does not have a unitary definition
[but, for an effort to ground open science in first principles, see 
@tennantValuePropositionOpen2020].
Instead, several guidelines exist for practices that are considered "open".
The TOP-guidelines are one of the most influential concrete operationalisations
of Open Science principles [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis; and 8) replication of published results.
WORCS defines the goals of open science in terms of these guidelines, and is
designed to facilitate meeting each of these guidelines, with one exception: We
do not address replication of published results, because replication relates to
the subject of a research project, not to its execution.
We group the remaining seven guidelines into three categories: citation (1),
sharing (2-5), and preregistration (6-7).

<!--## Goal of this paper

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations.
However, large organizations are slow to change.
For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that
they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
At the same time, <!-- edit -->
<!--The formation of "Open Science Communities" at universities around the globe
illustrates the momentum of this grassroots movement
[@brinkmanWidescaleAdoptionOpen2020]. -->

## Implementation in R

Although the principles underlying this workflow are universal, WORCS has first
been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
Several arguments support the choice to work in R.
First, R and all of it extensions are free open source software, which
should be the tool of choice for open science.
Second, as of this writing, R is the most advanced programming language in terms
of the implementation of tools required for an open science workflow.
Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
Finally, R is highly interoperable:
Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
Fortran, and Python [@allaireMarkdownPythonEngine]. There are excellent free
resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].

Working with R is simplified immensely by using the RStudio integrated
development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
RStudio automates and streamlines tedious or complicated aspects of working with
the tools used in WORCS, and many are embedded directly into the visual user
interface.
This makes RStudio a comprehensive solution for open science research projects.
Another important feature of RStudio is project management.
A project bundles writing, analyses, data, references, etcetera, into a
self-contained folder, that can be uploaded entirely to 'GitHub' and downloaded
by future users. The `worcs` R-package installs a new RStudio project template.
When a new project is initialized from this template, the bookkeeping required
to set up an open science project is performed automatically. The `worcs` 
package has a [vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)
on how to set up your system with all of the required tools.

@@

WORCS relies on several free, open source software solutions.
The first is *dynamic document generation* (DDG): Writing scientific reports in
a format that interleaves written reports with blocks of code used to conduct
the analyses. The text is automatically formatted as a scientific paper in 
several potential styles. When the text is formatted, the results of the code
blocks are
automatically generated and insterted in the text, or rendered as Figures and
Tables. Dynamic document generation supercedes
the classical approach of using separate programs to write prose and conduct
analyses, and then manually copy-pasting analysis results into the text.

@@

In the R-implementation of WORCS, we recommend centering a research project
around one dynamically generated 'RMarkdown' document, which includes all 
analysis code. Longer scripts can be
stored in `.R` files, and called from the main document using the `source()`
function. When a reader or reviewer compiles this document, all code is run
automatically, thus verifying computational replicability. The 'RMarkdown' 
document can be automatically formatted in many styles, including
APA style [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020],
a host of other scientific formats 
[see @allaireRticlesArticleFormats2020], and as plain 'Markdown' for 'GitHub'.

@@

WORCS recommends using the *explicit* version control software 'Git'
(www.git-scm.com). Git tracks
changes to files, and stores these changes when the user makes a "commit" (a 
snapshot of the version controlled files).
Git retains a complete log of all commits, and users can compare changes between
different commits, or go back to a previous version of the code (for example,
after making a mistake, or to replicate a previous version of the results).
A project version controlled with Git is called a "Repository", or Repo.
Git only version controls files explicitly committed by the user. Moreover, it
is possible to prevent files from being version controlled - which is useful for
privacy sensitive data.
A `.gitignore` file lists files that should not be version-controlled.
These files thus exist only on the user's private computer.

The functionality of 'Git' is amplified by services such as 'GitHub'
(https://github.com). 'GitHub' is best understood as a cloud storage service
with social networking functionality. The cloud storage aspect
of 'GitHub' works as follows: You can "clone" (copy) a local 'Git' repository to
the 'GitHub' website, as a backup or research archive. The social network aspect
comes into play when a repository on 'GitHub' is set to "public": This allows 
other researchers to peruse the repository and see how the work was done, clone
it to their own computer to replicate the original work or apply the methods to
their own data, open "Issues" to ask questions or give feedback on the project,
or even send a "Pull request" with suggested changes to the text or code for
your consideration. 'Git' and 'GitHub' shine as tools for collaboration, because
different people can simultaneously work on different parts of a project, and
their changes can be compared and automatically merged on the website. Even on
solo projects, working with 'Git'/'GitHub' has many benefits: Staying organized,
being able to start a new study with a clone of an old, similar repository, or
splitting off an "experimental branch" to try something new, while retaining the
ability to "revert" (return) to a previous state of the project, or to "merge"
(incorporate) the experimental branch.

The R-implementation of WORCS uses the package `gert`, a 'Git' client for 'R'
[@oomsGertSimpleGit2019], to connect a local project to a remote repository on
'GitHub'. The `worcs` package also contains a user-friendly function to set
the default user credentials, `git_user()`, and a function to add, commit, and
push all changed files in one step, `git_update()`. It is worth noting that,
although this paper focuses on 'GitHub', which is the most widely used 'Git'
cloud storage service, two notable alternative services exist: 
'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). As of
version 0.1.2, the `worcs` template dialog accepts 'https' addresses of 
repositories on both of these platforms interchangeably. The package has a
[vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) with
additional instructions for using these platforms. However,
`worcs` has been most extensively tested with 'GitHub'.

@@

<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc. This could be a place where readers stumble and stop reading. Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective. 
[main issue] it should become clear earlier what the WORCS workflow entails: e.g. switching to working with a particular R Project Template. Discuss the ‘workflow' section earlier. Related: Can you make a schematic overview of the WORCS?overview--> 
Many solutions exist to ensure computational reproducibility. These solutions typically work by
enveloping your research project in a distinct "environment" that only has access to programs that
are explicitly installed, and maintaining a record of these programs. These solutions differ in
user-friendliness and effectiveness. If strict reproducibility is required, users might choose to conduct all analyses in a 'Docker container': An environment that behaves like a
virtual computer, which can be stored like a sort of time capsule, and
identically reinstated on a user computer, or in the cloud.
This is effective, and also preserves software outside the R environment (e.g.,
preprocessing tools for neuroimaging data). An excellent Docker-based workflow
was recently introduced by @peikertReproducibleDataAnalysis2019. This workflow
is based on many of the same principles as WORCS, and the two approaches are
conceptually and practically compatible. However, the workflow described by 
Peikert and Brandmaier can be difficult to set up for novice users. An R-package
is currently in development to facilitate this transition [@repro]. Another 
user-friendly Docker-based workflow is offered by the cloud-based collaboration
platform 'Digital Ocean'. Although Digital Ocean enjoys some popularity among 
scientists, it is not clear whether this platform's
["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/)
is GDPR compliant - it seems to focus more on limiting liability than on
compliance. The necessity to upload data for analysis is a risk with any
cloud-computing solution when human participants are involved.

@@

A key consideration when developing a research project is what filetypes to use.
A case can be made to use text-based files exclusively (or as much as possible),
instead of binary files.
Text-based files can be read by machines and humans alike.
Binary files, such as 'Word' (`.docx`) or 'SPSS' (`.sav`, `.spo`) files,
must be decoded first, and many require commercial software to do so.
Text-based files are also required when using 'Git' for version control.
In 'Git', each change reflects a line of text added or removed. This means that, when
writing a version-controlled paper in plain-text, it is recommended to insert a
line break after every sentence (instead of after every paragraph). That way,
the change log will indicate which specific sentence was edited.
Binary files should not be version controlled with 'Git',
as changes in the encrypted file are uninterpretable. Binary files are also often 
larger than text-based files, which means they take up more space on cloud
hosting services, such as 'GitHub'. For these reasons, uploading large binary
files to these platforms is frowned upon.
'GitHub' encourages the use of text-based files by allowing visitors of their website to peruse
text-based files online, and
conveniently renders certain filetypes: For example, 'GitHub' displays `.md`
(Markdown) files as web pages, and `.csv` files as spreadsheets. This is why
`worcs` uses text-based file types like Markdown and `.csv` extensively.

@@

## Preparing your system

Before you can use the R-implementation of the WORCS workflow, you have to install the required software.
You only have to perform this setup procedure only once for each computer you
intend to use `worcs` on. It should take about 30 minutes, and it is documented
in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
The setup vignette does not assume a prior installation of R, so it is suitable for novice users.
After setting up your system, you can use the WORCS workflow for all of your
projects. This step-by-step procedure is detailed in [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).
In this paper, we provide only a brief outline of the WORCS procedure (as of Version `r worcs:::cran_version()`) below.
We highly recommend users to refer to [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html) for the most up-to-date procedure (subject to improvements).

@@

# How WORCS relates to the TOP-guidelines

@@

To facilitate citing datasets, researchers sometimes publish *data papers*;
documents that detail the procedure, sample, and codebook. Specialized journals,
such as the
[*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/),
aid in the publication of these data papers. For smaller projects, researchers
sometimes publish their data in an online repository - like 'GitHub' - along
with a text file with the preferred citation for the data (which can be a
substantive paper), and the license that applies to the data, such as Creative
Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or
[BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/). When in doubt,
one can always contact the data creators and ask what the preferred citation is.

References for software are sometimes provided within the software environment;
for example, in R, package references can be found by
calling `citation("packagename")`. This returns an APA-style reference, and a
BibTeX entry. Software papers are also common, and are sometimes called
"Application Notes". The [Journal of Open Source Software (JOSS)](https://joss.theoj.org/)
offers programmers a way to generate a publication 
based on well-documented software;
in this process, the code itself is peer reviewed, and writing an actual paper is optional. For an example, see
@rosenbergTidyLPAPackageEasily2018.
Software citations can also sometimes be found 
in the source code repository (e.g., on 'GitHub'), using the
['Citation File Format'](https://citation-file-format.github.io/).

@@

<!-- LB: As I reader, I would prefer to hear about data sharing options before the synthetic data part (which is only relevant to those who cannot share their data). E.g. can I indicate a point in my workflow (e.g. after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->
If data can be shared openly, they can simply be pushed to a public 'Git'
repository, along with the analysis code. This way, others can download the
entire repository and reproduce the analyses from start to finish. 
If data cannot be shared, researchers should aim to safeguard the potential for
computational reproducibility and secondary analysis as much as possible. WORCS
recommends two solutions to accomplish this goal.
The first solution is to publish a *checksum* of the original data file.
Think of a checksum as a 32-character summary, or as a "one word" description,
of the contents of a file.
Any change to the file will result in a different checksum. 
Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
When data cannot be shared, the risk of fraud or abuse can be mitigated by publishing a checksum for the original data, in addition to the complete analysis code.
Using the checksum to verify the identity of a private dataset, researchers can prove to an independent auditer that running the public analysis code on the private data results in the published results of their work.

@@

In `worcs`, researchers can call the function `open_data(data)` to make the data
publicly available. This function stores the `data` object (a `data.frame` or
`matrix`) in a spreadsheet file (called `data.csv` by default), and generates an
accompanying codebook that can be viewed online.
All changed files should now be (manually) added to the 'Git' repository, committed, and
pushed to GitHub. Once the GitHub repository is made public, these data will be
open to the public.
Assume that, once you do this, the data cannot be un-shared.

Alternatively, if the project requires data to remain closed, researchers can
call `closed_data(data)`. This function also stores the data in a local `.csv` file
and generates a codebook,
but the original data file is added to `.gitignore` so it cannot be
accidentally added to the 'Git' repository.
The function also computes a checksum for the original data, and logs it in the
`.worcs` project file. This checksum serves as the indelible record of the state 
of the raw data.
The function `closed_data()` also creates a synthetic dataset (called
`synthetic_data.csv` by default).
The user should push all changed files 'GitHub'.
Once the 'GitHub' repository is made public, people will have access to a
checksum for the original data that exist on your local machine, a codebook
describing those data, and a synthetic copy of the data.

@@

When writing a manuscript created using dynamic document generation, analysis
code is embedded in the prose of the paper. Thus, the TOP-guideline of sharing
analysis code can be met simply by committing the source code of the manuscript
to the 'Git' repository, and making this repository Public, e.g., on 'GitHub'.
If authors additionally use open data and a reproducible environment (as suggested
in WORCS), then a third party can simply replicate all analyses by copying the
entire repository from the 'Git' hosting service, and Knitting the manuscript on
their local computer.

>>>>

abstract: |
  Adopting open science
  principles can be challenging and costly, because doing so requires
  substantial conceptual education and training in the use of new tools. The 
  goal of this paper is to remove these obstacles. We introduce the Workflow for
  Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
  procedure that researchers can follow to make an entire project open and
  reproducible. It is based on best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies. To facilitate widespread adoption, the
  WORCS principles are implemented in the R-package `worcs`, which offers an
  RStudio 
  project template that is, effectively, a "one-click solution" for R users. 
  This paper introduces the workflow, discusses how it meets different 
  requirements of open science, and provides examples of the implementation of
  `worcs` in R.

@@

The (social) sciences are arguably past the tipping point of a paradigm shift
towards open science.
Several immediate causes rallied support for this transition, including
highly publicized cases of scientific fraud [as discussed by
@leveltFailingScienceFraudulent2012], increasing awareness of questionable
research practices and their consequences for the credibility of findings [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis
[@shroutPsychologyScienceKnowledge2018].
However, open science should not be seen as a cure (or punishment) for this
crisis. As the late dr. Jonathan Tennant put it: "Open science is just good science"
[@tennantOpenScienceJust2018].
Open science creates opportunities for researchers to more easily
conduct reliable, cumulative, and 
collaborative science [@nosekScientificUtopiaOpening2012;
@adolphOpenBehavioralScience2012]. Open science also
promotes inclusivity, because it removes barriers for participation.
Capitalizing on these advances has the potential to accellerate scientific
progress [see also @coyneReplicationInitiativesWill2016].

Many researchers are motivated to adopt best practices for open
science and enjoy these benefits. And yet, the question of "how" to 
apply open science practices can be daunting. Making the transition
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by presenting a simple workflow 
that meets most requirements open science, and is based on best practices:
The Worfkflow for Open Reproducible 
Code in Science (WORCS). WORCS is a lightweight approach to open science and
computational reproducibility. The workflow (Figure <!--\@ref(fig:workflow)-->) 
constitutes a step-by-step procedure that researchers can follow to make a 
research project open and reproducible. WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them. It can also be used in the absence of top-down support or guidelines
to produce work in accordance with best practices.
Although the principles underlying this workflow are universal, WORCS has first
been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
The R-package `worcs` offers a project template for RStudio
[@rstudioteamRStudioIntegratedDevelopment2015], which is, effectively,
a "one click solution". This paper introduces the WORCS workflow, discusses how
it meets best practices for open science, and illustrates the use of the `worcs`
package.

@@

Although Open Science is advocated by many, it does not have a unitary definition
[for an effort to ground open science in first principles, see 
@tennantValuePropositionOpen2020].
Instead, several guidelines exist for practices that are considered "open".
The TOP-guidelines are one of the most influential concrete operationalisations
of Open Science principles [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis; and 8) replication of published results.
WORCS defines the goals of open science in terms of these guidelines, and is
designed to facilitate meeting each of these guidelines, with one exception: We
do not address replication of published results, because replication relates to
the subject of a research project, not to its execution.
We group the remaining seven guidelines into three categories: citation (1),
sharing (2-5), and preregistration (6-7).

@@

WORCS relies on several free, open source software solutions.
The first is *dynamic document generation* (DDG): Writing scientific reports in
a format that interleaves written reports with blocks of code used to conduct
the analyses. The text is automatically formatted as a scientific paper in 
several potential styles. When the text is formatted, the code
blocks are
evaluated and their results are insterted in the text, or rendered as Figures and
Tables. Dynamic document generation supercedes
the classical approach of using separate programs to write prose and conduct
analyses, and then manually copy-pasting analysis results into the text.
The source code of this paper can serve as an example of DDG, and is
[available here](https://github.com/cjvanlissa/worcs/blob/master/paper/Manuscript.Rmd).

@@

In the R-implementation of WORCS, we recommend centering a research project
around one dynamically generated 'RMarkdown' document, which includes all 
analysis code. Longer scripts can be
stored in `.R` files, and called from the main document using the `source()`
function. When a reader or reviewer compiles this document, all code is run
automatically, thus verifying computational replicability. The 'RMarkdown' 
document can be automatically formatted in many styles, including
APA style [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020],
a host of other scientific formats 
[see @allaireRticlesArticleFormats2020], and as plain 'Markdown'.

@@

An integral part of WORCS is the *explicit* version control software 'Git'
(www.git-scm.com). Git tracks
changes to files, and stores these changes when the user makes a "commit" (a 
snapshot of the version controlled files).
Git retains a complete log of all commits, and users can compare changes between
different commits, or go back to a previous version of the code (for example,
after making a mistake, or to replicate a previous version of the results).
A project version controlled with Git is called a "Repository", or Repo.
Git only version controls files explicitly committed by the user. Moreover, it
is possible to prevent files from being version controlled - which is useful for
privacy sensitive data.
A `.gitignore` file lists files that should not be version-controlled.
These files thus exist only on the user's private computer.

The functionality of 'Git' is amplified by services such as 'GitHub'
(https://github.com). 'GitHub' is best understood as a cloud storage service
with social networking functionality. The cloud storage aspect
of 'GitHub' works as follows: You can "clone" (copy) a local 'Git' repository to
the 'GitHub' website, as a backup or research archive. The social network aspect
comes into play when a repository on 'GitHub' is set to "public": This allows 
other researchers to peruse the repository and see how the work was done, clone
it to their own computer to replicate the original work or apply the methods to
their own data, open "Issues" to ask questions or give feedback on the project,
or even send a "Pull request" (a request to incorporate some code)
with suggested changes to the text or code for
your consideration. 'Git' and 'GitHub' shine as tools for collaboration, because
different people can simultaneously work on different parts of a project, and
their changes can be compared and automatically merged on the website. Even on
solo projects, working with 'Git'/'GitHub' has many benefits: Staying organized,
being able to start a new study with a clone of an old, similar repository, or
splitting off an "experimental branch" to try something new, while retaining the
ability to "revert" (return) to a previous state of the project, or to "merge"
(incorporate) the experimental branch.

It is worth noting that,
although this paper focuses on the most widely used 'Git'
cloud storage service, 'GitHub', two notable alternatives exist: 
'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). As of
version 0.1.2, the `worcs` template dialog accepts 'https' addresses from 
repositories hosted on either of these platforms interchangeably.
The package has a
[vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) with
additional instructions for using alternative platforms. In 
order to present the workflow in a platform-agnostic way, we will refer to such
cloud hosting services as "remote repositories" in the remainder of this
paper.
However, `worcs` has been tested most extensively with 'GitHub'.

The R-implementation of WORCS uses the package `gert`, a 'Git' client for 'R'
[@oomsGertSimpleGit2019], to connect a local project to a remote repository.
The `worcs` package also contains a user-friendly function to set
the default user credentials, `git_user()`, and a function to add, commit, and
push all changed files to a remote repository in one step, `git_update()`.

@@

Many solutions exist to ensure computational reproducibility. These solutions typically work by
enveloping your research project in a distinct "environment" that only has access to programs that
are explicitly installed, and maintaining a record of these programs. These solutions differ in
user-friendliness and effectiveness. If strict reproducibility is required, users might choose to conduct all analyses in a 'Docker container': An environment that behaves like a
virtual computer, which can be stored like a sort of time capsule, and
identically reinstated on a user computer, or in the cloud.
This is effective, and also preserves software outside the R environment (e.g.,
preprocessing tools for neuroimaging data). An excellent Docker-based workflow
was recently introduced by @peikertReproducibleDataAnalysis2019. This workflow
is based on many of the same principles as WORCS, and the two approaches are
conceptually and practically compatible. However, the workflow described by 
Peikert and Brandmaier can be difficult to set up for novice users. An R-package
is currently in development to facilitate this transition [@repro]. Another 
user-friendly Docker-based workflow is offered by the cloud-based collaboration
platform 'Digital Ocean'. Although Digital Ocean enjoys some popularity among 
scientists, it is not clear whether this platform's
["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/)
is GDPR compliant - it seems to focus more on limiting liability than on
compliance. The necessity to upload data for analysis is a risk with any
cloud-computing solution when human participants are involved.

@@

A key consideration when developing a research project is what filetypes to use.
A case can be made to use text-based files exclusively (or as much as possible),
instead of binary files.
Text-based files can be read by machines and humans alike.
Binary files, such as 'Word' (`.docx`) or 'SPSS' (`.sav`, `.spo`) files,
must be decoded first, and many require commercial software to do so.
Text-based files are also required when using 'Git' for version control.
In 'Git', each change reflects a line of text added or removed. This means that, when
writing a version-controlled paper in plain-text, it is recommended to insert a
line break after every sentence (instead of after every paragraph). That way,
the change log will indicate which specific sentence was edited.
Binary files should not be version controlled with 'Git',
as changes in the encrypted file are uninterpretable. Binary files are also often 
larger than text-based files, which means they take up more space on cloud
hosting services. For these reasons, uploading large binary
files to remote repositories is frowned upon. 
WORCS encourages the use of
text-based files for all purposes.
It is worth noting that 'GitHub' also rewards the use of text-based files by
rendering certain text-based filetypes: 
For example, `.md`
(Markdown) files are displayed as web pages, and `.csv` files as spreadsheets.

@@

We provide a conceptual outline of the WORCS procedure below, based on WORCS
Version 0.1.2. 
For users of the `worcs` R-package, we highly recommend referring to
[the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
which is more detailed and updated with future developments.

@@

## The R-implementation of WORCS

The R-package `worcs` automate most steps of the workflow
for R users.
Several arguments support the choice to implement this workflow in R.
First, R and all of it extensions are free open source software, which
should be the tool of choice for open science.
Second, as of this writing, R is the most advanced programming language in terms
of the implementation of tools required for an open science workflow.
Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
Finally, R is highly interoperable:
Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
Fortran, and Python [@allaireMarkdownPythonEngine]. There are excellent free
resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].

Working with R is simplified immensely by using the RStudio integrated
development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
RStudio automates and streamlines tedious or complicated aspects of working with
the tools used in WORCS, and many are embedded directly into the visual user
interface.
This makes RStudio a comprehensive solution for open science research projects.
Another important feature of RStudio is project management.
A project bundles writing, analyses, data, references, etcetera, into a
self-contained folder, that can be uploaded entirely to a 'Git' remote repository and downloaded
by future users. The `worcs` R-package installs a new RStudio project template.
When a new project is initialized from this template, the bookkeeping required
to set up an open science project is performed automatically.

### Preparing your system

Before you can use the R-implementation of the WORCS workflow, you have to install the required software.
You only have to perform this setup procedure only once for each computer you
intend to use `worcs` on. It should take about 30 minutes, and it is documented
in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
The setup vignette does not assume a prior installation of R, so it is suitable for novice users.
After setting up your system, you can use `worcs` for all of your
projects. The R-implementation of the workflow introduced earlier is detailed,
step-by-step, in
[the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).

# How WORCS helps meet the TOP-guidelines

@@

To facilitate citing datasets, researchers sometimes publish *data papers*;
documents that detail the procedure, sample, and codebook. Specialized journals,
such as the
[*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/),
aid in the publication of these data papers. For smaller projects, researchers often
simply share the data in a remote repository, along
with a text file with the preferred citation for the data (which can be a
substantive paper), and the license that applies to the data, such as Creative
Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or
[BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/). When in doubt,
one can always contact the data creators and ask what the preferred citation is.

References for software are sometimes provided within the software environment;
for example, in R, package references can be found by
calling `citation("packagename")`. This returns an APA-style reference, and a
BibTeX entry. Software papers are also common, and are sometimes called
"Application Notes". The [Journal of Open Source Software (JOSS)](https://joss.theoj.org/)
offers programmers a way to generate a publication 
based on well-documented software;
in this process, the code itself is peer reviewed, and writing an actual paper is optional. For an example, see
@rosenbergTidyLPAPackageEasily2018.
Software citations can also sometimes be found 
in the source code repository (e.g., on 'Git' remote repositories), using the
['Citation File Format'](https://citation-file-format.github.io/).

@@

If data can be shared openly, they can simply be pushed to a public 'Git'
repository, along with the analysis code. This way, others can download the
entire repository and reproduce the analyses from start to finish. 
If data cannot be shared, researchers should aim to safeguard the potential for
computational reproducibility and secondary analysis as much as possible. WORCS
recommends two solutions to accomplish this goal.
The first solution is to publish a *checksum* of the original data file.
Think of a checksum as a 32-character summary, or as a "one word" description,
of the contents of a file.
Any change to the file will result in a different checksum. 
Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
When data cannot be shared, the risk of fraud or abuse can be mitigated by publishing a checksum for the original data, in addition to the complete analysis code.
Using the checksum to verify the identity of a private dataset, researchers can prove to an independent auditer that running the public analysis code on the private data results in the published results of their work.

@@

In `worcs`, researchers can call the function `open_data(data)` to make the data
publicly available. This function stores the `data` object (a `data.frame` or
`matrix`) in a spreadsheet file (called `data.csv` by default), and generates an
accompanying codebook that can be viewed online.
All changed files should now be (manually) added to the 'Git' repository, committed, and
pushed to the remote repository. Once the remote repository is made public, these data will be
open to the public.
Assume that, once you do this, the data cannot be un-shared.

Alternatively, if the project requires data to remain closed, researchers can
call `closed_data(data)`. This function also stores the data in a local `.csv` file
and generates a codebook,
but the original data file is added to `.gitignore` so it cannot be
accidentally added to the 'Git' repository.
The function also computes a checksum for the original data, and logs it in the
`.worcs` project file. This checksum serves as the indelible record of the state 
of the raw data.
The function `closed_data()` also creates a synthetic dataset (called
`synthetic_data.csv` by default).
The user should push all changed files to the remote repository.
Once the remote repository is made public, people will have access to a
checksum for the original data that exist on your local machine, a codebook
describing those data, and a synthetic copy of the data.

@@

When writing a manuscript created using dynamic document generation, analysis
code is embedded in the prose of the paper. Thus, the TOP-guideline of sharing
analysis code can be met simply by committing the source code of the manuscript
to the 'Git' repository, and making this remote repository Public.
If authors additionally use open data and a reproducible environment (as suggested
in WORCS), then a third party can simply replicate all analyses by copying the
entire repository from the 'Git' hosting service, and Knitting the manuscript on
their local computer.
